---
title: "DMAP Resources and Examples"
author: "Marc Weber"
format: revealjs
embed-resources: true
editor: visual
execute: 
  eval: true
  error: false
  echo: true
  warning: false
  include: true
---

## Data Commons and DMAP Portal

-   The best place to start exploring DMAP resources is [the Data Commons home page](https://work.epa.gov/data-science/data-commons) and [the DMAP Portal page](https://work.epa.gov/data-science/dmap-portal) with links to:

    -   current services
    -   tutorials
    -   showcases of current
    -   uses in the agency
    -   example code snippets

## Data Training Resources

Also look at the [training resources](https://work.epa.gov/data-science/dmap-training-resources) page where you can:

-   sign up for the DMAP newsletter
-   read FAQs
-   see a calendar of all upcoming data science related meetings in the agency

## Using the CLI for managing data transfers

-   You can use the Analytics Platform interface, but it will be slow for big data transfers and it's best to use the command line interface (CLI)
-   You'll need to install AWS CLI

## Using Sagemaker

-   Sagemaker is an AWS machine learning service that provides infrastructure, tools, workflows for machine learning
-   Sagemaker is set to replace the current Analytics platform in DMAP sometime this year

## Accessing Data Commons data in RStudio on the desktop

This is following example code [here](https://work.epa.gov/data-science/data-commons)

## OMS data bucket

```{r}
library(aws.s3)
bucket_contents <-
  data.table::rbindlist(get_bucket(bucket = "edap-oms-data-commons"),
  fill = TRUE)
head(bucket_contents$Key)
```

## ORD data bucket

```{r}
bucket_contents <-
  data.table::rbindlist(get_bucket(bucket = "dmap-data-commons-ord"),
  fill = TRUE)
head(bucket_contents$Key)
```

## My public content in ORD data bucket

Read in some of sample geoparquet file I have stored in DMAP ORD public bucket


## R code for accessing
```{r}
library(arrow);library(data.table);library(readr)
'NARS_Sample_Frames/CT_points.parquet' %in% bucket_contents$Key

test_data <- s3read_using(
  FUN = read_parquet,
  bucket = "dmap-data-commons-ord",
  object = "NARS_Sample_Frames/CT_points.parquet",
  show_col_types = TRUE,
  col_names=TRUE,
  guess_max = 10000
)

head(test_data)
```
