---
title: "Spatial Group 8/23/2023"
author: "Marc Weber"
format: revealjs
self-contained: true
editor: visual
---

## A Little Background

- Originally this meeting was between Ryan and I and GIS contractor starting back in 2015?

- Meeting has evolved over time - now an informal meeting time to discuss spatial topics

- [Teams space for the group](https://teams.microsoft.com/l/entity/com.microsoft.teamspace.tab.wiki/tab::425d8b6b-1959-4006-a9e6-41c5a91ddfd1?context=%7B%22subEntityId%22%3A%22%7B%5C%22pageId%5C%22%3A2%2C%5C%22origin%5C%22%3A2%7D%22%2C%22channelId%22%3A%2219%3A50c2782cfb5041afacdc66f8fec51f93%40thread.skype%22%7D&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)

## Topics We've Covered

- Earth Engine workshop - Marc
- R leaflet / mapview - Ryan / Marc
- tmap - Amalia / Marc
- R packages - Michael
- R StreamNetworkTools - Darin
- R nhdplusTools - Marc / Amalia
- Hands-on intro to git - Rick
- Story Map demo - Jessie
- Remote sensing libraries in R - Marc

## Topics We've Covered...

- Reprex - Michael
- Intro to DMAP - Andrew Chase
- Walk-through /  demo using DMAP - Marc
- Repetitive tasks in R (apply family) - Darin
- Using parallel in R - Michael
- geoconnex - Marc
- Web services exercise in R/Python - Marc
- Discussion of current work - Hannah
- Intro to Quarto - Marc

## Topics We've Covered...
- Alpine Lakes research - Amalia
- Using DMAP - Darrin Kopp
- Modifiable areal unit problem (MAUP) discussion / paper discussion
- Seasonal and spatial variation in stream nitrate loading - Amalia
- Qualitative social science and spatial analysis - Marissa
- Recap on ICRW8 R/Python spatial workshop - Marc
- All these are listed in [Teams Wiki page for group](https://teams.microsoft.com/l/entity/com.microsoft.teamspace.tab.wiki/tab::425d8b6b-1959-4006-a9e6-41c5a91ddfd1?context=%7B%22subEntityId%22%3A%22%7B%5C%22pageId%5C%22%3A2%2C%5C%22sectionId%5C%22%3A5%2C%5C%22origin%5C%22%3A2%7D%22%2C%22channelId%22%3A%2219%3A50c2782cfb5041afacdc66f8fec51f93%40thread.skype%22%7D&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)

## Future Ideas

- Table of future discussion topics [here](https://teams.microsoft.com/l/entity/com.microsoft.teamspace.tab.wiki/tab::425d8b6b-1959-4006-a9e6-41c5a91ddfd1?context=%7B%22subEntityId%22%3A%22%7B%5C%22pageId%5C%22%3A2%2C%5C%22sectionId%5C%22%3A4%2C%5C%22origin%5C%22%3A2%7D%22%2C%22channelId%22%3A%2219%3A50c2782cfb5041afacdc66f8fec51f93%40thread.skype%22%7D&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)
- Include things such as:
  - spmodel
  - another Earth Engine Workshop
  - Dask / parallelization in Python
  - QGIS Intro
  - Data Ecosystems
  - ScienceHub Geospatial
  - Python geemap, segment-geospatial (SAM model)
  
## Reminder of agency GIS resources and groups

- [ORD Geospatial Sciences CoOp](https://teams.microsoft.com/l/channel/19%3a1ee6a984ac0b4d32a7109bc638a85014%40thread.skype/ORD%2520Geospatial?groupId=93022b12-6005-468e-a542-501d13b4caa8&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)
- [EPA Geospatial Teams Channel](https://teams.microsoft.com/l/team/19%3a01dbde32e2da416e8f2ab76409521c0c%40thread.skype/conversations?groupId=8cb556ff-b4b3-4652-bfc7-b42285928d03&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7) (multiples grops / channels)
- [Agency Geospatial Landing Page](https://work.epa.gov/geospatial-resources)
  - Links to tutorials, SEGS, Policies, etc
- GeoPlatform Webinar Series / Training
  - Every Monday, contact Kaitlin Kean or Alex O'Neal
- [R CoP](https://teams.microsoft.com/l/channel/19%3aa6272ef8a55c404ea56fa19145076a3d%40thread.skype/R%2520User%2520Group?groupId=93022b12-6005-468e-a542-501d13b4caa8&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)
- [Python Cop](https://teams.microsoft.com/l/channel/19%3a7d92e05a84bf4cba9c31e38711304cd5%40thread.skype/Python?groupId=93022b12-6005-468e-a542-501d13b4caa8&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)

## Some new tools to introduce
- `QGIS` and `WhiteboxTools` are software / tool suites that offer additional open source GIS tools and functionality above and beyond what is available in COTS software like `ArcPRO`
- We can have upcoming sessions to do an intro or take a deeper dive into using either of these tools, but there are packages in R to link to both of these:
- `qgisprocess` and `whitebox` R packages

## `qgisprocess`
Just a demo of what is available at the `qgisprocess` [GitHub](https://github.com/r-spatial/qgisprocess) site

Run a buffer and dissolve operation on built-in `sf` North Carolina dataset using the native:buffer algorithm from `qgis`
```{r}
#| echo: true
#| code-fold: true
library(qgisprocess)
input <- sf::read_sf(system.file("shape/nc.shp", package = "sf"))
```

## North Carolina features before a buffer / dissolve
```{r}
#| echo: true
plot(input$geometry)
```

## Run algorithm
Note the class of the result of the algorithm
```{r}
#| echo: true
result <- qgis_run_algorithm(
  "native:buffer",
  INPUT = input,
  DISTANCE = 1,
  DISSOLVE = TRUE
)

result
```

## Plot the buffer and dissolve results
We can convert our result to an `sf` object easily
```{r}
#| echo: true
output_sf <- sf::st_as_sf(result)
plot(sf::st_geometry(output_sf))
```

## Getting help on a tool
```{r}
#| echo: true
qgisprocess::qgis_show_help("native:buffer")
```

## Tool search
```{r}
#| echo: true
qgisprocess::qgis_search_algorithms(algorithm = "overlap")
```

## Available tools
```{r}
#| echo: true
qgis_algorithms()
```