---
title: "Spatial Group Meeting"
date: '2023-08-23'
author: "Marc Weber"
format: revealjs
self-contained: true
editor: visual
---

## A Little Background

-   Originally this meeting was between Ryan and I and GIS contractor starting back in 2015?

-   Meeting has evolved over time - now an informal meeting time to discuss spatial topics

-   [Teams space for the group](https://teams.microsoft.com/l/entity/com.microsoft.teamspace.tab.wiki/tab::425d8b6b-1959-4006-a9e6-41c5a91ddfd1?context=%7B%22subEntityId%22%3A%22%7B%5C%22pageId%5C%22%3A2%2C%5C%22origin%5C%22%3A2%7D%22%2C%22channelId%22%3A%2219%3A50c2782cfb5041afacdc66f8fec51f93%40thread.skype%22%7D&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)

## Topics We've Covered

-   Earth Engine workshop - Marc
-   R leaflet / mapview - Ryan / Marc
-   tmap - Amalia / Marc
-   R packages - Michael
-   R StreamNetworkTools - Darin
-   R nhdplusTools - Marc / Amalia
-   Hands-on intro to git - Rick
-   Story Map demo - Jessie
-   Remote sensing libraries in R - Marc

## Topics We've Covered...

-   Reprex - Michael
-   Intro to DMAP - Andrew Chase
-   Walk-through / demo using DMAP - Marc
-   Repetitive tasks in R (apply family) - Darin
-   Using parallel in R - Michael
-   geoconnex - Marc
-   Web services exercise in R/Python - Marc
-   Discussion of current work - Hannah
-   Intro to Quarto - Marc

## Topics We've Covered...

-   Alpine Lakes research - Amalia
-   Using DMAP - Darrin Kopp
-   Modifiable areal unit problem (MAUP) discussion / paper discussion
-   Seasonal and spatial variation in stream nitrate loading - Amalia
-   Qualitative social science and spatial analysis - Marissa
-   Recap on ICRW8 R/Python spatial workshop - Marc
-   All these are listed in [Teams Wiki page for group](https://teams.microsoft.com/l/entity/com.microsoft.teamspace.tab.wiki/tab::425d8b6b-1959-4006-a9e6-41c5a91ddfd1?context=%7B%22subEntityId%22%3A%22%7B%5C%22pageId%5C%22%3A2%2C%5C%22sectionId%5C%22%3A5%2C%5C%22origin%5C%22%3A2%7D%22%2C%22channelId%22%3A%2219%3A50c2782cfb5041afacdc66f8fec51f93%40thread.skype%22%7D&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)

## Future Ideas

-   Table of future discussion topics [here](https://teams.microsoft.com/l/entity/com.microsoft.teamspace.tab.wiki/tab::425d8b6b-1959-4006-a9e6-41c5a91ddfd1?context=%7B%22subEntityId%22%3A%22%7B%5C%22pageId%5C%22%3A2%2C%5C%22sectionId%5C%22%3A4%2C%5C%22origin%5C%22%3A2%7D%22%2C%22channelId%22%3A%2219%3A50c2782cfb5041afacdc66f8fec51f93%40thread.skype%22%7D&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)
-   Include things such as:
    -   spmodel
    -   another Earth Engine Workshop
    -   Dask / parallelization in Python
    -   QGIS Intro
    -   Data Ecosystems
    -   ScienceHub Geospatial
    -   Python geemap, segment-geospatial (SAM model)

## Reminder of agency GIS resources and groups

-   [ORD Geospatial Sciences CoOp](https://teams.microsoft.com/l/channel/19%3a1ee6a984ac0b4d32a7109bc638a85014%40thread.skype/ORD%2520Geospatial?groupId=93022b12-6005-468e-a542-501d13b4caa8&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)
-   [EPA Geospatial Teams Channel](https://teams.microsoft.com/l/team/19%3a01dbde32e2da416e8f2ab76409521c0c%40thread.skype/conversations?groupId=8cb556ff-b4b3-4652-bfc7-b42285928d03&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7) (multiples grops / channels)
-   [Agency Geospatial Landing Page](https://work.epa.gov/geospatial-resources)
    -   Links to tutorials, SEGS, Policies, etc
-   GeoPlatform Webinar Series / Training
    -   Every Monday, contact Kaitlin Kean or Alex O'Neal
-   [R CoP](https://teams.microsoft.com/l/channel/19%3aa6272ef8a55c404ea56fa19145076a3d%40thread.skype/R%2520User%2520Group?groupId=93022b12-6005-468e-a542-501d13b4caa8&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)
-   [Python CoP](https://teams.microsoft.com/l/channel/19%3a7d92e05a84bf4cba9c31e38711304cd5%40thread.skype/Python?groupId=93022b12-6005-468e-a542-501d13b4caa8&tenantId=88b378b3-6748-4867-acf9-76aacbeca6a7)

## Some new tools to introduce

-   `QGIS` and `WhiteboxTools` are software / tool suites that offer additional open source GIS tools and functionality above and beyond what is available in COTS software like `ArcPRO`
-   We can have upcoming sessions to do an intro or take a deeper dive into using either of these tools, but there are packages in R to link to both of these:
    -   `qgisprocess` and `whitebox` R packages
    -   Need `QGIS` and `WhiteboxTools` installed as stand-alone applications on your computer for these R packages to work.

## `qgisprocess`

Just a demo of what is available at the `qgisprocess` [GitHub](https://github.com/r-spatial/qgisprocess) site

We'll run a buffer and dissolve operation on built-in `sf` North Carolina dataset using the `native:buffer` algorithm from `qgis`

```{r}
#| echo: true
library(qgisprocess)
input <- sf::read_sf(system.file("shape/nc.shp", package = "sf"))
```

## North Carolina features before a buffer / dissolve

```{r}
#| echo: true
plot(input$geometry)
```

## Run algorithm

Note the class of the result of the algorithm

```{r}
#| echo: true
result <- qgis_run_algorithm(
  "native:buffer",
  INPUT = input,
  DISTANCE = 1,
  DISSOLVE = TRUE
)

result
```

## Plot the buffer and dissolve results

We can convert our result to an `sf` object easily

```{r}
#| echo: true
output_sf <- sf::st_as_sf(result)
plot(sf::st_geometry(output_sf))
```

## Getting help on a tool

```{r}
#| echo: true
qgisprocess::qgis_show_help("native:buffer")
```

## Tool search

```{r}
#| echo: true
qgisprocess::qgis_search_algorithms(algorithm = "overlap")
```

## Available tools

```{r}
#| echo: true
qgis_algorithms()
```

## Whitebox

Also just a demo of what is outlined in the `WhiteboxR` [GitHub repository](https://github.com/opengeos/whiteboxR).

-   After you have `WhiteboxTools` installed

    -   install the R library

    -   run the wb_init() function to point to location of executable

```{r}
#| echo: true
# install.packages("whitebox")
library(whitebox)
wbt_init(exe_path = 'C:/Users/mweber/OneDrive - Environmental Protection Agency (EPA)/Profile/Downloads/WBT/whitebox_tools.exe')
```

## Generate stream network with Whitebox

Using sample data and couple tools available in WhiteboxTools we can generate a flow accumulation from a DEM

First get sample DEM

```{r}
#| echo: true
# Set input raster DEM file
dem <- sample_dem_data()
```

## Apply processing steps

1.  Apply feature-preserving smoothing (FeaturePreservingSmoothing)

2.  Breach depressions in a DEM (BreachDepressions)

3.  Calculate D-Infinity flow accumulation (DInfFlowAccumulation)

## Code

```{r}
#| echo: true
## 1. FeaturePreservingSmoothing
wbt_feature_preserving_smoothing(
  dem = dem,
  output = "./smoothed.tif",
  filter = 9
)

## 2. BreachDepressions
wbt_breach_depressions(dem = "./smoothed.tif", output = "./breached.tif")

## 3. DInfFlowAccumulation
wbt_d_inf_flow_accumulation(input = dem, output = "./flow_accum.tif")
```

## Plot results with `terra`

```{r}
#| echo: true
terra::plot(terra::rast("./flow_accum.tif"))
```

## Available WhiteBox tools

List all tools with 'flow' in the name

```{r}
#| echo: true
library(printr)
wbt_list_tools("flow")
```

## Get tool help

```{r}
#| echo: true
wbt_tool_help("LongestFlowpath")
```
