---
title: "Web Services in R"
author: "Marc Weber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    self-contained: true
    code-link: true
    code-fold: true
editor: visual
---

## Setup

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(sf)
library(tidyverse)
library(readr)
remotes::install_github("r-arcgis/arcgis", dependencies = TRUE)
library(mapview)
mapviewOptions(fgb=FALSE, basemaps="OpenStreetMap")
library(AOI)
```

## Arcgislayers package

```{r}
library(arcgis)
```

### Examples from [GitHub repo for package](https://github.com/R-ArcGIS/arcgislayers)
```{r}
furl <- "https://services.arcgis.com/P3ePLMYs2RVChkJx/ArcGIS/rest/services/USA_Counties_Generalized_Boundaries/FeatureServer/0"

county_fl <- arc_open(furl)

county_fl
```

```{r}
arc_select(county_fl)
```

```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))
test <- arc_select(
  county_fl,
  filter_geom = sf::st_bbox(nc[1,])
)
```

```{r}
list_fields(county_fl)
```

```{r}
arc_select(
  county_fl, 
  fields = c("state_abbr", "population"), 
  where = "population > 1000000"
)
```


### Test with a few layers Mark mentioned
Try [ECHO](https://echogeo.epa.gov/arcgis/rest/services/ECHO/Facilities/MapServer) - note that we need to drill down to a particular numberred layer as listed on this link to grab features - examples below
```{r}
furl <- "https://echogeo.epa.gov/arcgis/rest/services/ECHO/Facilities/MapServer/3"

rcra <- arc_open(furl)
rcra
```

#### Look at table to figure out field query options
```{r}
list_fields(rcra)
```

#### Try a fields based query
```{r}
rcra_pdx <- arc_select(
  rcra, 
  fields = c("GLOBALID", "RCR_NAME", "RCR_CITY","RCR_STATE","RCR_ZIP"), 
  where = "RCR_ZIP = '97209'"
)
```

```{r}
mapview(rcra_pdx)
```

#### Try a spatial query
```{r}
Benton <- aoi_get(state = "OR", county= "Benton")
rcra_benton <- arc_select(
  rcra,
  filter_geom = sf::st_bbox(Benton))
mapview(rcra_benton)
```

### EJ Screen layer - food access
```{r}
furl <- "https://raisz.epa.gov/arcgis/rest/services/ejscreen/ejscreen_v2023_with_as_cnmi_gu_vi/MapServer/10"

food_access <- arc_open(furl)
food_access
```
#### Look at table to figure out field query options
```{r}
list_fields(food_access)
```

```{r}
Multnomah <- aoi_get(state = "OR", county= "Multnomah")
food_access_multnomah <- arc_select(
  food_access,
  filter_geom = sf::st_bbox(Multnomah))
mapview(food_access_multnomah)
```